### 2013-12-12 scripts/evaluate-spikein.Rmd
Continuing work on assembly script
  Bugs fixed
  - mistake to remove `trimclip` reads - need these for expression quantification!!!

  Additions
  - merge chimeric and redundant transcript using CAP3
  - remove *in silico* contigs from A22 assembly
  - gene expression quantification using `sailfish`

## Check 'spike-in'

Evaluate whether any BLAST hits to 'spike-in' reads from Trinity transcriptome assembly using all reads, but *not* including *in silico spike-in reads`

```{r evaluate_assembly, eval=TRUE}
# load spike-in fasta file
spikein <- "../data/sim/known-sim.fasta"
spikeinfa <- readDNAStringSet(spikein)
length(names(spikeinfa))

# make BLAST database from spike-in fasta
system(paste("makeblastdb -dbtype nucl -in ", spikein, sep=""))

# BLAST A22 assembled transcripts (including spike-in reads) against database
transcripts <- "../results/trinity-full/Trinity.fasta"
system(paste("blastn -query ", transcripts, " -db ", spikein, " -outfmt 6 -out blast_full.txt",sep=""))

# evaluate
blastfull <- "blast_full.txt"
try(assembly.spikein.eval(spikein, blastfull))
```

No hits!