### 2015-10-01 ApTranscriptome_resample.Rmd
script for resampling to calculate false-positive rate

# script to calculate empirical false-positive rate

library(MASS)
library(plyr)
library(dplyr)

# load functions
source("scripts/RxNseq.R")

# load real data
load("ApTranscriptome_exp_data.Rda")
str(TPM.dt.sub)

# dataframe to collect results
D <- data.frame(
  resamp = 1:10,
  num_signif = rep(NA, 10))

for(i in 1:100) {
  RxNresample <- TPM.dt.sub %>%
    group_by(Transcript, colony) %>%
    select(TPM) %>%
    sample_frac() %>%
    mutate(val = c(0, 3.5, 10.5, 14, 17.5, 21, 24.5, 28, 31.5, 35, 38.5))

  # define model for RxN function
  model <-  "log(TPM+1) ~ colony + val + I(val^2) + colony:val + colony:I(val^2)"

  # calculate overall P value and R^2 for each transcript
  RxNpval_resamp <- ddply(RxNresample, .(Transcript), .inform= TRUE, modpFunc, .parallel = TRUE)

  # adjust FDR
  RxNpval_resamp$padj <- p.adjust(RxNpval_resamp$pval, method = "fdr")

  # number of transcripts with FDR < 0.05
  numsignif <- length(which(RxNpval_resamp$padj < 0.05))

  # add to dataframe
  D[i, "num_signif"] <- numsignif
  print(D)
  }

save(D, file = "resample.out")